Cosmology with Weak Lensing photometric surveys: constraints with redshift tomography of non--Gaussian statistics

### 2016-04-18 photoz_paper/paper.tex
added table with constraints

\begin{abstract}
Weak gravitational lensing is becoming a mature technique to constrain cosmological parameters, such as the dark energy equation of state $w$. When analyzing galaxy surveys, redshift information has proven to be a valuable addition to angular shear correlations. We forecast parameter constraints on the triplet $(\Omega_m,w,\sigma_8)$ for a LSST like galaxy photometric survey, using tomography of the shear-shear power spectrum, convergence peak counts and higher convergence moments. We find that the addition of peak counts and moments to the power spectrum can significantly improve the constraints on the $(\Omega_m,w)$ doublet. We also find that most the cosmological information is contained in the highest redshift galaxies, with redshift tomography further improving the $w$ constraints by only 15\% to 30\%. We evaluate the effect of uncorrected photometric redshift systematics on parameter constraints. We find that different statistics lead to different bias directions in parameter space, leaving a window open for self--calibration. 
\end{abstract}

@@

Figure \ref{fig:constraintsOm-w} shows the $1\sigma$ confidence contours on the $(\Omega_m,w)$ doublet calculated with equation (\ref{meth:parcovestimator}) after the PCA dimensionality reduction performed according to equation (\ref{meth:pcaprojection}), for a variety of choices $({\rm statistic},N_c)$. We also show the improvements on the confidence contours when combining different summary statistics after the corresponding dimensionality reductions have been performed.

@@

@@

The number of components to consider increases when including redshift information, and can reach $\sim 30$ for the power spectrum and moments and $\sim 40$ for the peak counts. Figure \ref{fig:nopca} clearly that most of the cosmological information is concentrated in the last redshift bin. The use of redshift tomography tightens the constraints by approximately 33\% for the power spectrum, 16\% for the peaks and 25\% for the moments.

@@

\section{Conclusions}
In this work we study cosmological parameter constraints forecasts for an LSST-like galaxy survey using the convergence power spectrum and a range of non--Gaussian statistics. We make use of redshift tomography to improve the constraints from their single redshift counterparts. We also investigate the effects of uncorrected photo-$z$ systematic effects on the inferred cosmology. We find that:
\begin{itemize}
	\item Principal Component analysis is a robust technique to keep the dimensionality of the parameter space under control and avoid the numerical pitfalls explained in \citep{Taylor12,DodelsonSchneider13,Taylor14} and more recently \citep{PetriVariance}. In particular we find that only a few components are necessary to characterize the cosmological information content in single redshift statistics, while more components are necessary when tomography is included. Nevertheless we find that the number of required components $N_c$ is significantly smaller than the full summary statistic space dimensionality before performing PCA.
	\item Most of the cosmological information is contained in high redshift galaxies, with tomography improving $w$ constraints by only a few $10\%$.   
	\item Peak counts and moments can significantly improve $(\Omega_m,w)$ parameter constraints when used in combination with the power spectrum.
	\item Uncorrected photo-$z$ systematics can bias parameter constraints obtained from the power spectrum and the moments, but in different parameter directions, leaving open windows for self-calibration
\end{itemize}

>>>>

\begin{abstract}
Weak gravitational lensing is becoming a mature technique to constrain cosmological parameters, such as the dark energy equation of state $w$. When analyzing galaxy surveys, redshift information has proven to be a valuable addition to angular shear correlations. We forecast parameter constraints on the triplet $(\Omega_m,w,\sigma_8)$ for a LSST like galaxy photometric survey, using tomography of the shear-shear power spectrum, convergence peak counts and higher convergence moments. We find that the addition of peak counts and moments to the power spectrum can significantly improve the constraints on the $(\Omega_m,w)$ doublet. We also find that most the cosmological information on $w$ is contained in the highest redshift galaxies, with redshift tomography further improving the constraints by only 15\% to 30\%. The improvement obtained with redshift tomography is much bigger when considering the constraints on $(\Omega_m,w)$ using the convergence power spectrum. We evaluate the effect of uncorrected photometric redshift systematics on parameter constraints. We find that different statistics lead to different bias directions in parameter space, leaving a window open for self--calibration. 
\end{abstract}

@@

Figure \ref{fig:constraintsOm-w} shows the $1\sigma$ confidence contours on the $(\Omega_m,w)$ doublet calculated with equation (\ref{meth:parcovestimator}) after the PCA dimensionality reduction performed according to equation (\ref{meth:pcaprojection}), for a variety of choices $({\rm statistic},N_c)$. We also show the improvements on the confidence contours when combining different summary statistics after the corresponding dimensionality reductions have been performed. The forecast constraints on the $(\Omega_m,w,\sigma_8)$ parameter space for a variety of summary statistics are summarized in Table \ref{tbl:constraints}.

@@

\begin{table*}
\begin{tabular}{c|c|c|c|c|}
\toprule
                        \textbf{Statistic} & $\Delta \Omega_m$ & $\Delta w$ & $\Delta \sigma_8$ & $\sqrt[3]{\rm Volume}$ $(\Omega_m,w,\sigma_8)$ \\ \hline \hline
\midrule
     Power spectrum ($\bar{z}_5$) &            0.0222 &     0.0286 &            0.0298 &                                       0.0087 \\
      Power spectrum (tomography) &            0.0038 &     0.0213 &            0.0060 &                                       0.0042 \\
              Peaks ($\bar{z}_5$) &            0.0049 &     0.0316 &            0.0050 &                                       0.0046 \\
               Peaks (tomography) &            0.0042 &     0.0271 &            0.0043 &                                       0.0050 \\
            Moments ($\bar{z}_5$) &            0.0027 &     0.0276 &            0.0026 &                                       0.0034 \\
             Moments (tomography) &            0.0020 &     0.0214 &            0.0020 &                                       0.0028 \\ \hline \hline
           Power spectrum + peaks &            0.0021 &     0.0153 &            0.0026 &                                       0.0030 \\ 
         Power spectrum + moments &            0.0016 &     0.0150 &            0.0019 &                                       0.0025 \\
 Power spectrum + peaks + moments &            0.0015 &     0.0121 &            0.0018 &                                       0.0023 \\ \hline
\bottomrule
\end{tabular}
\caption{Constraints on the $(\Omega_m,w,\sigma_8)$ parameter triplet using different summary statistics and redshift information. Each column $(\Delta \Omega_m,\Delta w,\Delta \sigma_8)$ contains the 1 $\sigma$ constraints on a particular parameter, marginalized over the other two. The last volume contains the volume of the $\sigma$ ellipsoid in $(\Omega_m,w,\sigma_8)$ space, calculated as $\sqrt{\det \bbh{\Sigma}}$.}
\label{tbl:constraints}
\end{table*}

@@

The number of components to consider increases when including redshift information, and can reach $\sim 30$ for the power spectrum and moments and $\sim 40$ for the peak counts. Figure \ref{fig:nopca} clearly that most of the cosmological information is concentrated in the last redshift bin. The use of redshift tomography tightens the $w$ constraints by approximately 33\% for the power spectrum, 16\% for the peaks and 25\% for the moments. Redshift tomography seems to play a vital role when constraining the $(\Omega_m,\sigma_8)$ doublet using the convergence power spectrum.

@@

\section{Conclusions}
In this work we study cosmological parameter constraints forecasts for an LSST-like galaxy survey using the convergence power spectrum and a range of non--Gaussian statistics. We make use of redshift tomography to improve the constraints from their single redshift counterparts. We also investigate the effects of uncorrected photo-$z$ systematic effects on the inferred cosmology. We find that:
\begin{itemize}
	\item Principal Component analysis is a robust technique to keep the dimensionality of the parameter space under control and avoid the numerical pitfalls explained in \citep{Taylor12,DodelsonSchneider13,Taylor14} and more recently \citep{PetriVariance}. In particular we find that only a few components are necessary to characterize the cosmological information content in single redshift statistics, while more components are necessary when tomography is included. Nevertheless we find that the number of required components $N_c$ is significantly smaller than the full summary statistic space dimensionality before performing PCA.
	\item Most of the cosmological information on $w$ is contained in high redshift galaxies, with tomography improving constraints by only a few $10\%$. The advantage using tomography is much bigger when constraining $(\Omega_m,\sigma_8)$ using the convergence power spectrum.    
	\item Peak counts and moments can significantly improve $(\Omega_m,w)$ parameter constraints when used in combination with the power spectrum.
	\item Uncorrected photo-$z$ systematics can bias parameter constraints obtained from the power spectrum and the moments, but in different parameter directions, leaving open windows for self-calibration
\end{itemize}
